---
title: 'Titanic Survival - Kaggle'
author: 'Tobias Ford'
date: '`r Sys.Date()`'
output:
  html_document:
    fig.captions: yes
    keep_md: yes
    toc: yes
  pdf_document:
    toc: yes
---

```{r set-options, echo=FALSE}
options(width=120)
```

# Titanic Survival - Kaggle

## Introduction

Provided initial training set of data (train.csv) containing passenger roster with some key details.

### Load Data

```{r}
train <- read.csv('../input/train.csv', stringsAsFactors = F)
test <- read.csv('../input/test.csv', stringsAsFactors = F)

str(train)
```

### Data Dictionary

```{r}
head(train,1)
```

* PassengerId
* Survived
* Pclass
* Name
* Sex
* Age
* SibSp
* Parch
* Ticket
* Fare
* Cabin
* Embarked

## First level of investigation

### Distribution Based on Sex
```{r}
table(train$Sex)
summary(train$Sex)
prop.table(table(train$Sex))


prop.table(table(train$Survived))

train$SurvivedBoolean <- as.logical(train$Survived)
train$SurvivedLabel[train$SurvivedBoolean == TRUE] <- 'Survived'
train$SurvivedLabel[train$SurvivedBoolean == FALSE] <- 'Died'

prop.table(table(train$Sex, train$SurvivedLabel))
prop.table(table(train$Sex, train$SurvivedLabel),1)

```

### Diving into Age

```{r}
summary(train$Age)

train$Child <- FALSE
train$Child[train$Age < 18] <- TRUE

table(train$Child)
table(train$Child, train$Survived)

aggregate(Survived ~ Child + Sex, data=train, FUN=sum)
aggregate(Survived ~ Child + Sex, data=train, FUN=length)
aggregate(Survived ~ Child + Sex, data=train, FUN=function(x) {sum(x)/length(x)})

```

### Diving into Fare

```{r}
train$Fare2 <- '30+'
train$Fare2[train$Fare < 30 & train$Fare >= 20] <- '20-30'
train$Fare2[train$Fare < 20 & train$Fare >= 10] <- '10-20'
train$Fare2[train$Fare < 10] <- '<10'

aggregate(Survived ~ Fare2 + Pclass + Sex, data=train, FUN=function(x) {sum(x)/length(x)})

```
Determined something meaningful here and added it to the submission.
Women in class 3 who paid $20 or more for a fare didn't fair well (pun intended).


### Using Decision Trees

```{r}
library(rpart)
fit <- rpart(Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked,
               data=train,
               method="class")

plot(fit)
text(fit)

library(rattle)
library(rpart.plot)
library(RColorBrewer)
fancyRpartPlot(fit)

Prediction <- predict(fit, test, type = "class")

```

## Submission

# first attempt



```{r}
test$Survived <- rep(0, 418)
test$Survived <- 0
test$Survived[test$Sex == 'female'] <- 1
test$Survived[test$Sex == 'female' & test$Pclass == 3 & test$Fare >= 20] <- 0

head(test,10)

submit <- data.frame(PassengerId = test$PassengerId, Survived = test$Survived)
submit <- data.frame(PassengerId = test$PassengerId, Survived = Prediction)
write.csv(submit, file = "theyallperish.csv", row.names=FALSE)

```